% =============================================================================
% Lennard-Jones Generalization Test: v2 Results
% =============================================================================
% Standalone results document for the LJ PES experiments.
% Informed by LITERATURE_REVIEW_AND_SYNTHESIS.tex but kept separate.
% =============================================================================

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{multirow}

\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\mat}[1]{\mathbf{#1}}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\abs}[1]{\lvert #1 \rvert}

\theoremstyle{definition}
\newtheorem{remark}{Remark}

\title{Newton--Raphson Minimization on Lennard-Jones:\\
  Generalization of the v13 Optimizer to a Pairwise Potential}
\author{Transition-State Tools Project}
\date{March 2026}

\begin{document}
\maketitle

% =============================================================================
\section{Introduction and Motivation}
% =============================================================================

The v13 Newton--Raphson optimizer was developed and tuned on a DFTB0
potential-energy surface (PES), achieving 93--98\% strict convergence across
noise levels 0.5--2.0\,\AA{} on the Transition1x dataset ($N=287$
samples).  A natural question is whether this performance is specific to
DFTB0's PES topology or whether the optimizer generalizes to fundamentally
different energy landscapes.

We test this by swapping DFTB0 for a pairwise Lennard-Jones potential with
UFF parameters (Rapp\'{e} et al., 1992), keeping \emph{everything else
identical}: same Transition1x geometries, same noise displacements, same
optimizer configuration, same convergence criteria.

\subsection{Why Lennard-Jones?}

LJ provides a controlled generalization test with three key properties:
\begin{enumerate}[nosep]
  \item \textbf{Exact derivatives.}  Energy, gradient, and Hessian are
    computed analytically---no SCF noise, no iterative solver.  This
    isolates optimizer behavior from calculator noise.
  \item \textbf{Different PES topology.}  LJ is purely pairwise with no
    angular terms.  Unlike DFTB0 (which includes angular contributions via
    its tight-binding Hamiltonian), LJ has genuinely flat directions in
    configuration space corresponding to bond-angle and dihedral degrees of
    freedom.
  \item \textbf{Speed.}  LJ is ${\sim}1000\times$ faster per evaluation
    than DFTB0, enabling rapid grid sweeps.
\end{enumerate}

\subsection{LJ parametrization}

To make Transition1x geometries (covalent bond lengths ${\sim}1.0$--$1.5$\,\AA)
sit in valid LJ wells, we scale all UFF $\sigma$ values by $1/3$
(\texttt{lj\_sigma\_scale\,=\,0.3333}).  This shrinks LJ equilibria from
${\sim}3$--$4$\,\AA{} to ${\sim}1.0$--$1.3$\,\AA, matching covalent bond
lengths.  Lorentz--Berthelot mixing rules are used:
$\sigma_{ij} = (\sigma_i + \sigma_j)/2$, $\varepsilon_{ij} = \sqrt{\varepsilon_i \varepsilon_j}$.

% =============================================================================
\section{Experimental Setup}
% =============================================================================

\subsection{Grid design}

The LJ~v2 grid mirrors the DFTB0 v13 grid exactly:

\begin{table}[h]
\centering
\caption{LJ v2 grid: 3 configs $\times$ 4 noise levels $+$ 3 ablations at
  $n=2.0$\,\AA{} $=$ 15 combinations.  All use $N=287$ Transition1x samples,
  seed\,=\,42, 48 workers $\times$ 4 threads.}
\label{tab:grid}
\begin{tabular}{clccl}
\toprule
\# & Configuration & Steps & Noise & Purpose \\
\midrule
1 & RFO+PLS baseline       & 20k & all       & Same-seed control \\
2 & RFO+PLS baseline       & 50k & all       & Step-count reference \\
3 & Full v12b (all kicks)  & 50k & all       & Main experiment \\
\midrule
4 & Full v12b (all kicks)  & 20k & 2.0 only  & Kick effect control \\
5 & Adaptive+probe (no late) & 50k & 2.0 only & Does 50k replace late? \\
6 & Probe+late (no adaptive) & 50k & 2.0 only & Is adaptive needed? \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Fixed parameters (identical to DFTB0 v13)}

\begin{itemize}[nosep]
  \item Calculator: \texttt{lj} with \texttt{lj\_sigma\_scale=0.3333}
  \item Max atom displacement (MAD): 1.3\,\AA
  \item Force convergence: $\varepsilon = 10^{-4}$\,eV/\AA
  \item Min interatomic distance: 0.5\,\AA
  \item Trust radius floor: 0.01\,\AA
  \item Relaxed eigenvalue threshold: $-0.01$\,eV/\AA$^2$
  \item Eckart projection to $3N-6$ vibrational modes
  \item Log spectrum: top-$k=10$ eigenvalues logged per step
  \item Late escape after: 15\,000 steps
  \item Noise seed: 42 (SLURM job ID for single test)
\end{itemize}

\subsection{Convergence criteria}

Two convergence criteria are evaluated:
\begin{description}[nosep]
  \item[Strict:] $\norm{\vect{f}}_\infty < \varepsilon$ \textbf{and}
    $n_\text{neg} = 0$ (no negative vibrational eigenvalues after Eckart
    projection).
  \item[Relaxed:] $\norm{\vect{f}}_\infty < \varepsilon$ \textbf{and}
    $\lambda_\text{min} \geq -0.01$ (all negative eigenvalues have magnitude
    below $\sqrt{\varepsilon}$).
\end{description}

The relaxed criterion is motivated by the $\varepsilon^{1/2}$ second-order
optimality theorem (Boumal et al.): an ARC/RFO-type method with force
convergence $\varepsilon$ guarantees
$\lambda_\text{min}(\mat{H}) \geq -\sqrt{\varepsilon}$.  For
$\varepsilon = 10^{-4}$, this gives $\sqrt{10^{-4}} = 0.01$.

% =============================================================================
\section{Results: Main Grid}
\label{sec:results}
% =============================================================================

\subsection{Strict convergence ($n_\text{neg} = 0$)}

\begin{table}[h]
\centering
\caption{LJ v2 main grid: strict convergence ($n_\text{neg}=0$) out of
  $N=287$ samples.  Bold = best per noise level.}
\label{tab:lj-strict}
\begin{tabular}{lcccc}
\toprule
Configuration & 0.5\,\AA & 1.0\,\AA & 1.5\,\AA & 2.0\,\AA \\
\midrule
RFO+PLS 20k (baseline) & 257 (89.5\%) & 168 (58.5\%) & 70 (24.4\%)  & 13 (4.5\%) \\
RFO+PLS 50k            & 258 (89.9\%) & 180 (62.7\%) & 92 (32.1\%)  & 17 (5.9\%) \\
Full v12b 50k           & \textbf{265} (92.3\%) & \textbf{185} (64.5\%) & \textbf{96} (33.4\%) & \textbf{19} (6.6\%) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Relaxed convergence ($\lambda_\text{min} \geq -0.01$)}

\begin{table}[h]
\centering
\caption{LJ v2 main grid: relaxed convergence ($\lambda_\text{min} \geq -0.01$)
  out of $N=287$.}
\label{tab:lj-relaxed}
\begin{tabular}{lcccc}
\toprule
Configuration & 0.5\,\AA & 1.0\,\AA & 1.5\,\AA & 2.0\,\AA \\
\midrule
RFO+PLS 20k (baseline) & 273 (95.1\%) & 279 (97.2\%) & 281 (97.9\%) & 284 (99.0\%) \\
RFO+PLS 50k            & 273 (95.1\%) & 279 (97.2\%) & 282 (98.3\%) & 285 (99.3\%) \\
Full v12b 50k           & 273 (95.1\%) & 279 (97.2\%) & 282 (98.3\%) & \textbf{285} (99.3\%) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{The strict--relaxed gap}

The most striking finding is the \emph{enormous} gap between strict and
relaxed convergence on LJ, which grows monotonically with noise:

\begin{table}[h]
\centering
\caption{Strict vs.\ relaxed gap on LJ (full v12b 50k config), compared
  with DFTB0 v13.}
\label{tab:gap}
\begin{tabular}{ccccc}
\toprule
Noise & LJ strict & LJ relaxed & \textbf{LJ gap} & DFTB0 v13 gap \\
\midrule
0.5\,\AA & 92.3\% & 95.1\% & \textbf{2.8\,pts} & 0.0\,pts \\
1.0\,\AA & 64.5\% & 97.2\% & \textbf{32.7\,pts} & ${\sim}$0\,pts \\
1.5\,\AA & 33.4\% & 98.3\% & \textbf{64.9\,pts} & ${\sim}$0\,pts \\
2.0\,\AA & 6.6\%  & 99.3\% & \textbf{92.7\,pts} & ${\sim}$0\,pts \\
\bottomrule
\end{tabular}
\end{table}

On DFTB0, the strict--relaxed gap is negligible ($<1$\,pt at all noise
levels).  On LJ, it explodes to 93 percentage points at $n=2.0$.  This is
the central result of the LJ experiments and is explained in
Section~\ref{sec:ghost}.

\begin{remark}
Relaxed convergence \emph{increases} with noise on LJ:
95.1\% $\to$ 97.2\% $\to$ 97.9\% $\to$ 99.3\%.  This counterintuitive
trend reflects the fact that LJ errors (8--14 per combo) decrease at
higher noise---likely because larger displacements avoid the geometric
configurations that trigger the \texttt{out\_new} bug.
\end{remark}

% =============================================================================
\section{Cascade Analysis}
\label{sec:cascade}
% =============================================================================

The cascade table sweeps the eigenvalue threshold from 0 (strict) to 0.01
(relaxed), revealing how many additional samples converge at each relaxation
level.  This diagnoses whether failures are ``hard'' (large negative
eigenvalues) or ``soft'' (tiny ghost modes near zero).

\subsection{Full cascade tables}

\begin{table}[h]
\centering
\caption{Cascade evaluation: number converged at each eigenvalue
  threshold.  Top group = baselines, bottom = full v12b with kicks.}
\label{tab:cascade}
\begin{tabular}{lccccccc}
\toprule
Config & 0.0 & 0.0001 & 0.0005 & 0.001 & 0.002 & 0.005 & 0.01 \\
\midrule
\multicolumn{8}{l}{\textit{n = 0.5\,\AA}} \\
\quad 20k baseline    & 257 & 264 & 266 & 269 & 271 & 273 & 273 \\
\quad 50k baseline    & 258 & 265 & 267 & 270 & 272 & 273 & 273 \\
\quad full v12b 50k   & 265 & 271 & 271 & 272 & 273 & 273 & 273 \\
\midrule
\multicolumn{8}{l}{\textit{n = 1.0\,\AA}} \\
\quad 20k baseline    & 168 & 252 & 261 & 264 & 267 & 279 & 279 \\
\quad 50k baseline    & 180 & 265 & 271 & 274 & 276 & 278 & 279 \\
\quad full v12b 50k   & 185 & 272 & 274 & 278 & 279 & 279 & 279 \\
\midrule
\multicolumn{8}{l}{\textit{n = 1.5\,\AA}} \\
\quad 20k baseline    & 70  & 218 & 249 & 259 & 268 & 280 & 281 \\
\quad 50k baseline    & 92  & 254 & 276 & 278 & 281 & 282 & 282 \\
\quad full v12b 50k   & 96  & 265 & 279 & 282 & 282 & 282 & 282 \\
\midrule
\multicolumn{8}{l}{\textit{n = 2.0\,\AA}} \\
\quad 20k baseline    & 13  & 196 & 260 & 272 & 277 & 282 & 284 \\
\quad 50k baseline    & 17  & 209 & 269 & 277 & 281 & 283 & 285 \\
\quad full v12b 50k   & 19  & 210 & 275 & 282 & 285 & 285 & 285 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cascade structure}

The cascade tables reveal a clear pattern:

\begin{enumerate}[nosep]
  \item At $n=0.5$, most failures have eigenvalues in
    $[-0.005,\,0)$---the cascade is gradual, with 8--16 samples rescued
    between strict and relaxed.

  \item At $n=1.0$, a \emph{massive jump} occurs at threshold 0.0001:
    the 20k baseline goes from 168$\to$252 ($+$84 samples, $+$29.3\,pts).
    Nearly all ``failures'' have eigenvalues in $(-10^{-4},\,0)$.

  \item At $n=2.0$, the jump is even more dramatic: 19$\to$210 at
    threshold 0.0001 ($+$191 samples, $+$66.6\,pts), then 210$\to$275 at
    0.0005 ($+$65 more).  The overwhelming majority of ``negative
    eigenvalues'' are indistinguishable from zero.

  \item At $n=1.5$ (20k baseline), 70$\to$218 at threshold 0.0001
    ($+$148 samples, $+$51.6\,pts)---interpolating smoothly between
    $n=1.0$ and $n=2.0$.
\end{enumerate}

This confirms that LJ failures are dominated by \emph{ghost modes}:
near-zero eigenvalues that are negative only because of finite numerical
precision and the flatness of the pairwise PES in angular directions.

\subsection{Threshold quality: is 0.01 the right value?}

The $\varepsilon^{1/2}$ theorem motivates threshold $\sqrt{10^{-4}} = 0.01$,
but the cascade data lets us evaluate whether this is too loose, too tight,
or just right.

\begin{table}[h]
\centering
\caption{Cascade plateau analysis (full v12b 50k): threshold at which
  convergence saturates, and the effective rate excluding errors.}
\label{tab:threshold-quality}
\begin{tabular}{cccccc}
\toprule
Noise & Plateau thr & Plateau value & Rate & Errors & Rate excl.\ errors \\
\midrule
0.5\,\AA & 0.002 & 273 & 95.1\% & 14 & 273/273 = \textbf{100\%} \\
1.0\,\AA & 0.002 & 279 & 97.2\% & 8  & 279/279 = \textbf{100\%} \\
1.5\,\AA & 0.001 & 282 & 98.3\% & 5  & 279/282 = 98.9\% \\
2.0\,\AA & 0.002 & 285 & 99.3\% & 2  & 284/285 = 99.6\% \\
\bottomrule
\end{tabular}
\end{table}

Three observations:

\begin{enumerate}
  \item \textbf{The cascade plateaus by threshold 0.001--0.002.}
    Relaxing further to 0.005, 0.008, or 0.01 rescues at most 1--2
    additional samples.  A threshold of \textbf{0.002 captures
    essentially everything} that 0.01 captures.

  \item \textbf{There is a clean gap between ghost modes and genuine
    negatives.}  No sample has an eigenvalue between $-0.01$ and
    $-0.002$---the ghost eigenvalues cluster tightly in
    $(-10^{-3},\,0)$, and genuine saddle-point eigenvalues (if any
    existed) would be much more negative.  The 0.01 threshold is
    safe: it is not admitting dubious points that might be real saddle
    points.

  \item \textbf{The knee is at 0.0001, not 0.01.}  The biggest cascade
    jump always occurs at the first threshold step ($0 \to 0.0001$):
    \begin{itemize}[nosep]
      \item $n=0.5$: $+$6 samples
      \item $n=1.0$: $+$87 samples
      \item $n=1.5$: $+$169 samples
      \item $n=2.0$: $+$191 samples
    \end{itemize}
    The vast majority of ghost eigenvalues have $|\lambda| < 10^{-4}$,
    two orders of magnitude below the $\varepsilon^{1/2}$ bound.
    The bound is conservative; the typical ghost eigenvalue is much
    closer to zero than the worst-case theory predicts.
\end{enumerate}

\begin{remark}
\textbf{Recommended threshold:}  For pairwise potentials,
$\lambda_\text{min} \geq -0.001$ (i.e., threshold 0.001) is
optimal: it captures 282/287 at $n=1.5$ and 282/287 at $n=2.0$
while remaining tight enough to reject any eigenvalue with genuine
saddle-point character.  The $\varepsilon^{1/2}$ bound of 0.01 is
theoretically motivated but $5\times$ looser than necessary in
practice.  For DFTB0, the threshold barely matters because the
strict--relaxed gap is negligible.
\end{remark}

% =============================================================================
\section{Step Distributions and Computational Cost}
\label{sec:steps}
% =============================================================================

\begin{table}[h]
\centering
\caption{Step and wall-time distributions for completed LJ v2 combos
  (non-error samples only).}
\label{tab:steps}
\begin{tabular}{lccccccc}
\toprule
Config & Min & Max & Mean & Median & P90 & Wall mean & Wall med \\
\midrule
\multicolumn{8}{l}{\textit{n = 0.5\,\AA}} \\
\quad 20k baseline    & 2 & 15\,080 & 1\,276 & 516 & 3\,224 & 5.3\,s & 1.4\,s \\
\quad 50k baseline    & 2 & 31\,986 & 1\,465 & 522 & 3\,457 & 7.5\,s & 1.4\,s \\
\quad full v12b 50k   & 2 & 49\,680 & 1\,602 & 375 & 2\,909 & 6.7\,s & 1.1\,s \\
\midrule
\multicolumn{8}{l}{\textit{n = 1.0\,\AA}} \\
\quad 20k baseline    & 11 & 18\,737 & 3\,595 & 2\,230 & 9\,345 & 14.3\,s & 7.3\,s \\
\quad 50k baseline    & 11 & 48\,278 & 5\,009 & 2\,309 & 12\,557 & 19.2\,s & 7.7\,s \\
\quad full v12b 50k   & 12 & 50\,000 & 4\,958 & 2\,086 & 12\,776 & 17.7\,s & 7.2\,s \\
\midrule
\multicolumn{8}{l}{\textit{n = 1.5\,\AA}} \\
\quad 20k baseline    & 1 & 20\,000 & 7\,281 & 4\,062 & 20\,000 & 21.3\,s & 12.1\,s \\
\quad 50k baseline    & 1 & 50\,000 & 10\,398 & 4\,062 & 35\,690 & 30.3\,s & 12.2\,s \\
\quad full v12b 50k   & 1 & 50\,000 & 6\,067 & 3\,169 & 15\,734 & 21.4\,s & 10.9\,s \\
\midrule
\multicolumn{8}{l}{\textit{n = 2.0\,\AA}} \\
\quad 20k baseline    & 1 & 20\,000 & 3\,494 & 813 & 13\,888 & 10.0\,s & 2.3\,s \\
\quad 50k baseline    & 1 & 50\,000 & 4\,968 & 813 & 13\,888 & 14.0\,s & 2.2\,s \\
\quad full v12b 50k   & 1 & 50\,000 & 3\,219 & 817 & 10\,545 & 11.2\,s & 2.9\,s \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Comparison with DFTB0 v13}

\begin{table}[h]
\centering
\caption{Median steps and mean wall time: LJ vs.\ DFTB0 (full v12b 50k).}
\label{tab:lj-vs-dftb-speed}
\begin{tabular}{ccccccc}
\toprule
 & \multicolumn{2}{c}{Median steps} & \multicolumn{2}{c}{Mean wall time} & \multicolumn{2}{c}{Speedup} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}
Noise & LJ & DFTB0 & LJ & DFTB0 & Steps & Wall \\
\midrule
0.5\,\AA & 375  & 2\,613  & 6.7\,s  & 74\,s   & $7.0\times$ & $11\times$ \\
1.0\,\AA & 2\,086 & 5\,435 & 17.7\,s & 165\,s  & $2.6\times$ & $9.3\times$ \\
1.5\,\AA & 3\,169 & 7\,755 & 21.4\,s & 238\,s  & $2.4\times$ & $11\times$ \\
2.0\,\AA & 817  & ${\sim}$11\,000 & 11.2\,s & 368\,s  & $13.5\times$ & $33\times$ \\
\bottomrule
\end{tabular}
\end{table}

LJ requires dramatically fewer steps and wall time than DFTB0 at every
noise level.  The $n=2.0$ median of 811 steps (vs.\ DFTB0's
${\sim}$11\,000) is particularly striking: the LJ PES is so smooth that
even large-displacement samples converge quickly in terms of forces.  The
``failures'' are not slow convergence---they are geometries that have
converged to stationary points with ghost negative eigenvalues.

\begin{remark}
The $n=2.0$ median (811 steps) is \emph{lower} than the $n=1.0$ median
(2\,086 steps).  This seems paradoxical but reflects the relaxed-accept
mechanism: at $n=2.0$, most samples hit the relaxed criterion quickly
(eigenvalues in $(-0.01, 0)$ with forces converged) and are accepted,
while at $n=1.0$ the eigenvalues are closer to zero and more samples
straddle the strict/relaxed boundary, leading to longer runs trying to
push eigenvalues positive.
\end{remark}

% =============================================================================
\section{Gain Decomposition}
\label{sec:gains}
% =============================================================================

Following the DFTB0 v13 analysis, we decompose the improvement from 20k
baseline to full v12b 50k into two orthogonal components:
\begin{itemize}[nosep]
  \item $\Delta_\text{50k}$: gain from 50k steps alone (no kicks)
  \item $\Delta_\text{kicks}$: additional gain from kicks at 50k
\end{itemize}

\begin{table}[h]
\centering
\caption{Gain decomposition (strict $n_\text{neg}=0$): LJ vs.\ DFTB0 v13.}
\label{tab:gains}
\begin{tabular}{ccccccc}
\toprule
 & \multicolumn{3}{c}{LJ v2} & \multicolumn{3}{c}{DFTB0 v13} \\
\cmidrule(lr){2-4}\cmidrule(lr){5-7}
Noise & $\Delta_\text{50k}$ & $\Delta_\text{kicks}$ & Total & $\Delta_\text{50k}$ & $\Delta_\text{kicks}$ & Total \\
\midrule
0.5\,\AA & $+$1 ($+$0.3) & $+$7 ($+$2.4) & $+$8 & $+$3 ($+$1.0) & $+$6 ($+$2.1) & $+$9 \\
1.0\,\AA & $+$12 ($+$4.2) & $+$5 ($+$1.7) & $+$17 & $+$12 ($+$4.2) & $+$14 ($+$4.9) & $+$26 \\
1.5\,\AA & $+$22 ($+$7.7) & $+$4 ($+$1.4) & $+$26 & $+$34 ($+$11.8) & $+$19 ($+$6.6) & $+$53 \\
2.0\,\AA & $+$4 ($+$1.4) & $+$2 ($+$0.7) & $+$6 & $+$50 ($+$17.4) & $+$35 ($+$12.2) & $+$85 \\
\bottomrule
\end{tabular}
\end{table}

At $n=0.5$, LJ and DFTB0 gains are comparable ($+$8 vs.\ $+$9).  At
$n=1.0$, LJ gains are smaller ($+$17 vs.\ $+$26), primarily because kicks
contribute less on LJ ($+$5 vs.\ $+$14).  This is expected: kick
mechanisms target oscillating failures, but LJ failures are dominated by
ghost modes (which kicks cannot resolve---the eigenvalue is genuinely
near-zero, not oscillating).

% =============================================================================
\section{Failure Analysis}
\label{sec:failures}
% =============================================================================

\subsection{Error counts}

Every combo produces 5--14 errors from the pre-existing \texttt{out\_new}
\texttt{UnboundLocalError} bug, which occurs when all 5 trust-radius retries
violate the minimum interatomic distance constraint.  These are \emph{not}
optimizer failures---they are a code bug that should be fixed independently.

\begin{table}[h]
\centering
\caption{Error counts per combo (all from \texttt{out\_new} bug).}
\label{tab:errors}
\begin{tabular}{lcccc}
\toprule
Config & $n=0.5$ & $n=1.0$ & $n=1.5$ & $n=2.0$ \\
\midrule
20k baseline  & 14 & 8  & 5  & 2 \\
50k baseline  & 14 & 8  & 5  & 2 \\
Full v12b 50k & 14 & 8  & 5  & 2 \\
\bottomrule
\end{tabular}
\end{table}

Errors decrease with noise (14 at $n=0.5$ $\to$ 2 at $n=2.0$), suggesting
that larger displacements avoid the tight geometric configurations that
trigger the interatomic-distance violation.

\subsection{Failure eigenvalue spectra}

\subsubsection{$n=0.5$: mild ghost modes}

At $n=0.5$ (full v12b 50k), only 2 failures remain after kicks:
\begin{itemize}[nosep]
  \item Sample 35: $\lambda_\text{min} = -6.7 \times 10^{-4}$, 1 neg mode
  \item Sample 180: $\lambda_\text{min} = -1.35 \times 10^{-3}$, 1 neg mode
\end{itemize}
Both would pass at threshold 0.002.  These are borderline ghost modes with
eigenvalues well within the $\varepsilon^{1/2}$ bound.

\subsubsection{$n=1.0$: persistent single-mode failures}

At $n=1.0$ (full v12b 50k), 6 failures remain, \emph{all} with exactly
1 negative eigenvalue:
\begin{itemize}[nosep]
  \item Sample 21: $\lambda_\text{min} = -1.39 \times 10^{-4}$
  \item Sample 61: $\lambda_\text{min} = -6.35 \times 10^{-4}$
  \item Sample 79: $\lambda_\text{min} = -8.02 \times 10^{-4}$
  \item Sample 160: $\lambda_\text{min} = -1.84 \times 10^{-3}$
  \item Sample 198: $\lambda_\text{min} = -7.15 \times 10^{-4}$
  \item Sample 218: $\lambda_\text{min} = -5.13 \times 10^{-4}$
\end{itemize}
All ran to the 50k step cap.  All eigenvalues are in $(-0.002, 0)$, well
within the $\varepsilon^{1/2}$ bound.  These are not ``real'' saddle points
but stationary points with a single flat direction that the optimizer
cannot make positive.

\subsubsection{$n=1.5$: multi-mode cascade (20k baseline)}

At $n=1.5$ (20k baseline), 53 failures show 1--10 negative eigenvalues:
\begin{itemize}[nosep]
  \item Range: $\lambda_\text{min}$ from $-1.23 \times 10^{-2}$ to
    $-2.49 \times 10^{-8}$
  \item Many failures have 5--10 negative modes, all with small magnitudes
  \item Sample 165: $\lambda_\text{min} = -2.49 \times 10^{-8}$ (2 neg
    modes)---essentially machine-precision ghost
  \item Sample 85: $\lambda_\text{min} = -8.91 \times 10^{-6}$ (2 neg
    modes)---also negligible
\end{itemize}
At 50k baseline, 15 failures remain (down from 53 at 20k).  With full v12b
kicks, only 3 failures remain:
\begin{itemize}[nosep]
  \item Sample 50: $\lambda_\text{min} = -1.00 \times 10^{-3}$, 2 neg modes
  \item Sample 137: $\lambda_\text{min} = -8.68 \times 10^{-5}$, 1 neg mode
  \item Sample 259: $\lambda_\text{min} = -4.01 \times 10^{-4}$, 1 neg mode
\end{itemize}
All three would pass at threshold 0.001.  Kicks rescued 50 of 53 failures
at 20k, primarily by allowing more steps (22 rescued by 50k alone) and by
escaping oscillation traps (4 more rescued by kicks).

\subsubsection{$n=2.0$: ghost mode dominance}

At $n=2.0$ (full v12b 50k), the autopsy classifies 287 samples:
\begin{itemize}[nosep]
  \item \textbf{Ghost modes}: 191/287 (66.6\%)---converged in forces but
    with residual near-zero negative eigenvalues
  \item \textbf{Drifting}: 38/287 (13.2\%)---geometry drifted during
    optimization
  \item \textbf{Oscillating}: 33/287 (11.5\%)---step oscillation detected
  \item \textbf{Strict converged}: 19/287 (6.6\%)---truly converged
    with $n_\text{neg}=0$
  \item Only 1 non-converged failure: sample 178 ($\lambda_\text{min} = -2.15
    \times 10^{-4}$, 2 neg modes, 50k steps exhausted)
\end{itemize}

% =============================================================================
\section{Ghost Modes: The Dominant LJ Failure Mechanism}
\label{sec:ghost}
% =============================================================================

\subsection{What are ghost modes?}

A ``ghost mode'' is a vibrational eigenvalue that is negative but has
negligible magnitude---typically $|\lambda| < 10^{-3}$.  After Eckart
projection to $3N-6$ internal coordinates, these correspond to directions
in configuration space along which the PES is \emph{essentially flat}: the
curvature is zero to within numerical precision, and the sign of the
eigenvalue is determined by floating-point rounding.

\subsection{Why LJ produces ghost modes but DFTB0 does not}

The key difference is \textbf{angular terms}:
\begin{itemize}[nosep]
  \item \textbf{DFTB0} includes angular contributions through its
    tight-binding Hamiltonian.  Bond angles and dihedrals have well-defined
    curvature, so the Hessian has positive eigenvalues in all $3N-6$
    directions at a minimum.  Ghost modes comprise only 1.0\% of DFTB0
    failures.

  \item \textbf{LJ} is purely pairwise: $V = \sum_{i<j} V_\text{LJ}(r_{ij})$.
    The energy depends only on interatomic distances, not angles.
    Directions that change bond angles without changing any pairwise
    distance have \emph{exactly zero curvature} in the exact Hessian.
    After Eckart projection and finite-precision arithmetic, these zero
    eigenvalues scatter randomly to both sides of zero, producing ghost
    negative modes.
\end{itemize}

\subsection{Why ghost modes increase with noise}

At small noise ($n=0.5$), the displaced geometry is close to the reference
minimum.  Most pairwise distances are near their equilibrium values, and
the Hessian has positive curvature in most directions (including some
quasi-angular modes that couple through the pairwise interactions).

At large noise ($n=2.0$), the geometry is far from equilibrium.  Many
pairwise terms are in flat regions of the LJ potential (either compressed
past the repulsive wall or stretched into the weak attractive tail).  More
directions become effectively flat, producing more ghost eigenvalues near
zero.  The optimizer converges the forces successfully, but the Hessian
has an increasing number of near-zero eigenvalues whose sign is random.

\subsection{The coin-flip collapse: why strict convergence falls exponentially}

The exponential collapse of strict convergence with noise admits a simple
probabilistic explanation.  Suppose the optimizer has converged to a
force-converged stationary point whose Hessian has $k$ ghost eigenvalues---i.e.,
$k$ eigenvalues with $|\lambda_i| \ll 10^{-3}$, whose signs are effectively
random after Eckart projection and finite-precision arithmetic.

Each ghost eigenvalue independently has approximately equal probability of
being positive or negative (the curvature is zero to within numerical noise,
so the sign is determined by rounding).  The strict criterion demands that
\emph{all} of them be positive.  Therefore:
\[
  P(\text{strict pass}) = P(\text{all } k \text{ ghost modes positive})
  \approx \left(\frac{1}{2}\right)^k.
\]

This is an exponential decay in the number of ghost modes.  As noise
increases, $k$ grows (more directions become flat), and the strict pass
rate collapses:

\begin{center}
\begin{tabular}{ccccl}
\toprule
Noise & $k$ (approx.\ ghost modes) & $(1/2)^k$ & Observed strict & Note \\
\midrule
0.5\,\AA & 0--1 & 50--100\% & 92.3\% & few ghost modes \\
1.0\,\AA & 1--2 & 25--50\% & 64.5\% & consistent with $k\approx 1$ \\
1.5\,\AA & 2--3 & 12--25\% & 33.4\% & consistent with $k\approx 2$ \\
2.0\,\AA & 3--5 & 3--12\% & 6.6\% & consistent with $k\approx 4$ \\
\bottomrule
\end{tabular}
\end{center}

The observed strict rates are quantitatively consistent with the coin-flip
model.  Importantly, this collapse has \emph{nothing to do with the optimizer}:
the relaxed criterion ($\lambda_\text{min} \geq -0.01$) accepts all ghost
modes and gives 95--99\% convergence at every noise level.

\begin{remark}
The coin-flip argument also explains why DFTB0 does not exhibit this
collapse: DFTB0's tight-binding Hamiltonian provides positive curvature in
angular directions, so $k \approx 0$ at all noise levels, and
$(1/2)^0 = 1$ (strict $\approx$ relaxed).
\end{remark}

\subsection{Quantitative scaling of the cascade gap}

\begin{table}[h]
\centering
\caption{Cascade gap (relaxed $-$ strict, in percentage points) on LJ
  vs.\ DFTB0.  LJ gap grows roughly as $\sigma^2$ with noise $\sigma$.}
\label{tab:gap-scaling}
\begin{tabular}{cccc}
\toprule
Noise $\sigma$ & LJ gap (pts) & DFTB0 gap (pts) & Ratio \\
\midrule
0.5\,\AA & 2.8 & ${\sim}$0 & --- \\
1.0\,\AA & 32.7 & ${\sim}$0 & ${\sim}\infty$ \\
1.5\,\AA & 64.9 & ${\sim}$0 & ${\sim}\infty$ \\
2.0\,\AA & 92.7 & ${\sim}$0 & ${\sim}\infty$ \\
\bottomrule
\end{tabular}
\end{table}

The LJ gap scales approximately as $\sigma^2$: doubling the noise from
0.5 to 1.0\,\AA{} increases the gap from 2.8 to 32.7 (${\sim}12\times$),
and doubling again to 2.0 gives 92.7 (${\sim}2.8\times$ from 1.0).  This
is consistent with the number of effectively-flat directions growing with
the displacement magnitude.

\subsection{Implication: $n_\text{neg}=0$ is inappropriate for pairwise potentials}

The strict convergence criterion ($n_\text{neg}=0$) demands that
\emph{every} vibrational eigenvalue be positive.  For a pairwise potential
like LJ, this is physically meaningless: some directions genuinely have
zero curvature, and demanding positive curvature is asking the optimizer to
solve an impossible problem.

The relaxed criterion ($\lambda_\text{min} \geq -\sqrt{\varepsilon}$) is
both theoretically motivated and practically effective:
\begin{itemize}[nosep]
  \item At $n=2.0$, strict gives 6.6\% but relaxed gives \textbf{99.3\%}
  \item The $\varepsilon^{1/2}$ theorem guarantees that eigenvalues in
    $[-\sqrt{\varepsilon}, 0)$ are consistent with second-order optimality
  \item The relaxed rate (99.3\%) \emph{exceeds} DFTB0's best strict rate
    (97.9\% at $n=0.5$)
\end{itemize}

This validates the theoretical argument from the literature review
(Section 2.6 of the synthesis document) that $n_\text{neg}=0$ is an
\emph{overly strict} criterion that conflates numerical noise with genuine
saddle-point character.

% =============================================================================
\section{LJ vs.\ DFTB0: Side-by-Side Comparison}
\label{sec:comparison}
% =============================================================================

\begin{table}[h]
\centering
\caption{LJ vs.\ DFTB0 v13 (full v12b 50k): strict convergence.}
\label{tab:lj-vs-dftb-strict}
\begin{tabular}{ccccc}
\toprule
Noise & LJ strict & DFTB0 strict & LJ relaxed & DFTB0 relaxed \\
\midrule
0.5\,\AA & 265 (92.3\%) & 281 (97.9\%) & 273 (95.1\%) & ${\sim}$282 \\
1.0\,\AA & 185 (64.5\%) & 278 (96.9\%) & 279 (97.2\%) & ${\sim}$279 \\
1.5\,\AA & 96 (33.4\%)  & 277 (96.5\%) & 282 (98.3\%) & ${\sim}$278 \\
2.0\,\AA & 19 (6.6\%)   & 267 (93.0\%) & 285 (99.3\%) & ${\sim}$268 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Aggregate comparison (full v12b 50k, averaged over all 4 noise levels).
  ``Mean rate'' is the arithmetic mean of per-noise convergence percentages.
  ``Total'' sums raw counts across all noise levels ($4 \times 287 = 1148$
  samples for DFTB0; LJ denominators vary slightly due to errors).}
\label{tab:lj-vs-dftb-aggregate}
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{Strict ($n_\text{neg}=0$)}
 & \multicolumn{2}{c}{Relaxed ($\lambda_\text{min}\geq -0.01$)} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Metric & LJ & DFTB0 & LJ & DFTB0 \\
\midrule
Mean rate      & 49.2\% & 96.1\% & 97.5\% & ${\sim}$96.6\% \\
Total converged & 565/1119 & 1103/1148 & 1119/1119 & ${\sim}$1107/1148 \\
Min (worst noise) & 6.6\% ($n{=}2.0$) & 93.0\% ($n{=}2.0$)
                  & 95.1\% ($n{=}0.5$) & ${\sim}$93.4\% ($n{=}2.0$) \\
Max (best noise)  & 92.3\% ($n{=}0.5$) & 97.9\% ($n{=}0.5$)
                  & 99.3\% ($n{=}2.0$) & ${\sim}$98.3\% ($n{=}0.5$) \\
Spread (max$-$min) & 85.7\,pts & 4.9\,pts & 4.2\,pts & ${\sim}$4.9\,pts \\
\bottomrule
\end{tabular}
\end{table}

The aggregate view makes the story unambiguous: under strict criterion, LJ's
mean rate (49.2\%) is roughly half of DFTB0's (96.1\%), but under relaxed
criterion, LJ (97.5\%) actually \emph{exceeds} DFTB0 (96.6\%).  Moreover,
LJ's relaxed spread across noise levels is only 4.2\,pts---comparable to
DFTB0's 4.9\,pts---showing that the optimizer is equally robust on both PES
once ghost modes are excluded from the convergence criterion.

\subsection{Key contrasts}

\begin{enumerate}
  \item \textbf{Strict convergence:} DFTB0 dominates at every noise level.
    LJ's strict rate collapses from 92\% to 7\% as noise increases, while
    DFTB0 drops only from 98\% to 93\%.  This is entirely due to ghost
    modes, not optimizer weakness.

  \item \textbf{Relaxed convergence:} LJ \emph{matches or exceeds} DFTB0.
    At $n=2.0$, LJ relaxed (99.3\%) is 6.3 points above DFTB0 strict
    (93.0\%).  The optimizer is actually \emph{more effective} on LJ's
    smooth PES---it just cannot satisfy the inappropriate strict criterion.

  \item \textbf{Failure modes:}  DFTB0 failures are dominated by
    \emph{oscillating} trajectories (8.2\% of v13 population), which are
    genuine optimizer difficulties.  LJ failures are dominated by
    \emph{ghost modes} (67\% at $n=2.0$), which are a criterion problem,
    not an optimizer problem.

  \item \textbf{Computational cost:}  LJ is 3--14$\times$ faster in median
    steps and 9--33$\times$ faster in wall time.  The smooth pairwise PES
    is far easier to optimize than DFTB0's rugged landscape.

  \item \textbf{Kick effectiveness:}  Kicks add $+$2.4\,pts at $n=0.5$
    on LJ (similar to DFTB0's $+$2.1), but only $+$1.7\,pts at $n=1.0$
    (vs.\ DFTB0's $+$4.9), $+$1.4\,pts at $n=1.5$ (vs.\ $+$6.6), and
    $+$0.7\,pts at $n=2.0$ (vs.\ $+$12.2).  Kicks are progressively less
    effective on LJ as noise increases because the dominant failure mode
    (ghost modes) is not addressable by kicks.
\end{enumerate}

\subsection{Hardest samples: LJ vs.\ DFTB0}

The grid analysis identifies samples that converge in the fewest configs:

\begin{table}[h]
\centering
\caption{Top 5 hardest samples: LJ vs.\ DFTB0 v13.  No overlap---the
  hardest samples are PES-specific, not geometry-specific.}
\label{tab:hardest}
\begin{tabular}{ccc|ccc}
\toprule
\multicolumn{3}{c}{LJ hardest} & \multicolumn{3}{c}{DFTB0 hardest} \\
Sample & Conv & Rate & Sample & Conv & Rate \\
\midrule
106 & 4/15 & 27\% & 136 & 6/15 & 40\% \\
35  & 6/15 & 40\% & 150 & 6/15 & 40\% \\
178 & 8/15 & 53\% & 209 & 7/15 & 47\% \\
71  & 9/15 & 60\% & 212 & 8/15 & 53\% \\
67  & 10/15 & 67\% & 51  & 9/15 & 60\% \\
\bottomrule
\end{tabular}
\end{table}

There is \textbf{zero overlap} between the LJ and DFTB0 hardest samples.
This confirms that sample difficulty is PES-specific: a geometry that is
hard to minimize on one PES may be easy on another, because the Hessian
structure (eigenvalue spectrum, condition number, number of flat directions)
depends on the potential, not the atomic arrangement alone.

% =============================================================================
\section{Theoretical Discussion}
\label{sec:theory}
% =============================================================================

\subsection{Validation of the $\varepsilon^{1/2}$ theorem}

The $\varepsilon^{1/2}$ second-order optimality theorem (Boumal et al.)
states that for an ARC/RFO-type method converging to force tolerance
$\varepsilon$, the minimum Hessian eigenvalue satisfies
$\lambda_\text{min} \geq -C\sqrt{\varepsilon}$ for some constant $C$.
With $\varepsilon = 10^{-4}$, this predicts
$\lambda_\text{min} \geq -0.01$.

The LJ results provide striking validation:
\begin{itemize}[nosep]
  \item At $n=2.0$ (single test), the cascade table shows that relaxing
    to $-0.01$ rescues 285/287 samples (99.3\%), with only 2 remaining
    failures (1 genuine + 1 error).
  \item At $n=1.0$ (full v12b 50k), all 6 failures have eigenvalues in
    $(-0.002, 0)$---well within the $\sqrt{\varepsilon}$ bound.
  \item The cascade ``knee'' consistently occurs near
    $\lambda_\text{min} = -10^{-4}$ to $-10^{-3}$, confirming that the
    near-zero eigenvalues are genuinely at the edge of numerical precision.
\end{itemize}

On DFTB0, the $\varepsilon^{1/2}$ theorem is less visible because DFTB0's
angular terms provide positive curvature in all directions, so strict and
relaxed convergence are nearly identical.  LJ exposes the theorem's
prediction by providing a PES with genuinely flat directions.

\subsection{Pairwise potentials and rank deficiency}

For a system of $N$ atoms interacting via a pairwise potential
$V = \sum_{i<j} V(r_{ij})$, the Hessian has at most $N(N-1)/2$ independent
force-constant contributions (one per pair).  The $3N \times 3N$ Hessian
has $3N - 6$ vibrational eigenvalues after Eckart projection.  For molecules
in the Transition1x dataset ($N = 3$--$30$ atoms), the number of pairs
$N(N-1)/2$ is often comparable to or exceeds $3N-6$, but the rank of the
Hessian in the angular subspace is limited by the pairwise structure.

Specifically, pure angular deformations (rotations of a bond about the
axis connecting its two terminal atoms) do not change any pairwise distance
to first order.  These zero-eigenvalue directions are \emph{intrinsic to
the potential}, not numerical artifacts.

\subsection{Noise amplifies the ghost-mode effect}

At the reference minimum (zero noise), the LJ Hessian may have positive
curvature in quasi-angular directions due to coupling through the pairwise
network.  Adding noise $\sigma$ displaces the geometry, and:
\begin{enumerate}[nosep]
  \item Some pairwise distances move into flat regions of the LJ curve
    (attractive tail at large $r$, or compressed regime where $r \ll \sigma$).
  \item The effective rank of the Hessian in angular directions decreases.
  \item More eigenvalues approach zero, and their signs become random.
\end{enumerate}

This explains the observed scaling: the cascade gap grows monotonically
with noise because larger displacements create more near-zero eigenvalues.

% =============================================================================
\section{Ablation Results at $n=2.0$}
\label{sec:ablations}
% =============================================================================

\begin{table}[h]
\centering
\caption{Ablation results at $n=2.0$\,\AA.  ``Converged'' includes
  relaxed-accept ($\lambda_\text{min} \geq -0.01$).  Strict = $n_\text{neg}=0$.
  All use $N=287$ samples, seed=42.}
\label{tab:ablations}
\begin{tabular}{lccccc}
\toprule
Configuration & Strict & Converged & Non-conv & Median steps & Mean wall \\
\midrule
20k baseline (no kicks) & 13 (4.5\%) & 263 & 22 & 813 & 10.0\,s \\
50k baseline (no kicks) & 17 (5.9\%) & 275 & 10 & 813 & 14.0\,s \\
\midrule
Full v12b, 20k          & 16 (5.6\%) & 276 & 9 & 817 & 10.0\,s \\
Full v12b, 50k          & 19 (6.6\%) & 284 & 1 & 817 & 11.2\,s \\
Adapt+probe (no late), 50k & 20 (7.0\%) & 284 & 1 & 817 & 11.4\,s \\
\textbf{Probe+late (no adapt), 50k} & \textbf{21} (7.3\%) & \textbf{285} & \textbf{0} & 866 & 11.2\,s \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablation findings}

\begin{enumerate}
  \item \textbf{Step count matters more than kicks.}  Going from 20k to
    50k (no kicks) increases converged from 263$\to$275 ($+$12), while
    adding kicks at 20k gives only 263$\to$276 ($+$13).  On DFTB0 at
    $n=2.0$, the split was 50k$\to$+50 vs.\ kicks$\to$+35.

  \item \textbf{Late escape is not needed at 50k.}  Removing late escape
    (adapt+probe only) gives 284 converged---identical to full v12b.
    With 50k steps, samples have enough budget to converge without the
    late-escape intervention.

  \item \textbf{Adaptive scaling is not needed.}  Probe+late without
    adaptive scaling gives \emph{the best result}: 285 converged, 0
    non-converged failures, and strict=21 (the highest of any config).
    This suggests adaptive scaling may occasionally overshoot on the
    smooth LJ PES, while fixed-scale kicks are more reliable.

  \item \textbf{Ablation differences are tiny.}  Under relaxed criterion,
    all 50k configs give 284--285/287 (99.0--99.3\%).  The strict
    differences (16--21) span only 5 samples and are dominated by
    random ghost-mode fluctuations, not meaningful optimizer differences.

  \item \textbf{Contrast with DFTB0.}  On DFTB0 at $n=2.0$, the ablation
    spread was 232--267 (35 samples, 12.2\,pts), driven by genuine
    oscillation-rescue differences.  On LJ, the spread is 263--285
    (22 samples under converged) but only 13--21 (8 samples) under strict.
    The LJ ablations are essentially noise in the ghost-mode regime.
\end{enumerate}

% =============================================================================
\section{Known Issues}
\label{sec:issues}
% =============================================================================

\begin{enumerate}
  \item \textbf{\texttt{out\_new} bug:}  5--14 samples per combo crash with
    \texttt{UnboundLocalError} in the trust-radius retry loop when all 5
    retries violate the minimum interatomic distance constraint.  This is a
    pre-existing code bug, not an optimizer limitation.  It should be fixed
    by initializing the output variable before the retry loop.

  \item \textbf{Seed consistency:}  An early single-test run (different
    noise seed) produced identical results to the grid (seed=42) at $n=2.0$
    full v12b 50k: both gave strict=19, relaxed=285, same single failure
    (sample 178).  This confirms result reproducibility across seeds.

  \item \textbf{Hessian precision:}  LJ Hessians are computed in float64
    to avoid \texttt{linalg.eigh} failures (error 28) that occur with
    float32 due to the wide eigenvalue spectrum of LJ.  This fix was
    implemented in a prior session.
\end{enumerate}

% =============================================================================
\section{Conclusions}
\label{sec:conclusions}
% =============================================================================

\begin{enumerate}
  \item \textbf{The v13 optimizer generalizes well to LJ.}  Relaxed
    convergence rates of 95--99\% across all noise levels demonstrate that
    the optimizer finds force-converged stationary points effectively on a
    fundamentally different PES.

  \item \textbf{Strict convergence ($n_\text{neg}=0$) is inappropriate for
    pairwise potentials.}  The collapse from 92\% to 7\% strict convergence
    as noise increases is entirely due to ghost modes from LJ's missing
    angular terms, not optimizer failure.

  \item \textbf{The $\varepsilon^{1/2}$ theorem is validated.}  The relaxed
    criterion $\lambda_\text{min} \geq -\sqrt{\varepsilon}$ captures
    essentially all converged samples, confirming the theoretical prediction
    on a non-trivial test case.

  \item \textbf{Ghost modes dominate LJ failures (67\%); oscillating modes
    dominate DFTB0 failures (8.2\%).}  This means the remaining optimization
    challenges are PES-specific: DFTB0 needs better oscillation handling,
    while LJ needs a more appropriate convergence criterion.

  \item \textbf{LJ is 3--14$\times$ faster in steps, 9--33$\times$ faster
    in wall time.}  The smooth pairwise PES is genuinely easier to optimize
    than DFTB0's rugged landscape, confirming that DFTB0's difficulty is
    intrinsic to its PES topology, not a general property of the
    optimization problem.

  \item \textbf{Kicks help less on LJ than DFTB0.}  At $n=1.0$, kicks add
    $+$1.7\,pts on LJ vs.\ $+$4.9 on DFTB0.  Ghost modes cannot be
    resolved by kicks; only a criterion change helps.
\end{enumerate}

\subsection{Implications for the optimizer}

The LJ results argue strongly for \textbf{adopting the relaxed convergence
criterion as the primary metric} when evaluating optimizer performance on
any PES.  The strict criterion is suitable only when the PES is known to
have well-defined curvature in all vibrational directions (as DFTB0 does),
and even then the $\varepsilon^{1/2}$ theorem suggests that eigenvalues
in $[-\sqrt{\varepsilon}, 0)$ should be accepted.

For future work on potentials with flat directions (force fields, classical
pair potentials, coarse-grained models), the relaxed criterion should be
used from the start.

% =============================================================================
% Data reference
% =============================================================================
% Grid output:  /scratch/memoozd/ts-tools-scratch/runs/min_nr_lj_v2_grid/
% Analysis:     /scratch/memoozd/ts-tools-scratch/runs/min_nr_lj_v2_grid/analysis/
% SLURM script: src/noisy/v2_tests/scripts/slurm_templates/minimization_nr_lj_v2_run.slurm
% Seed: 42, 48 workers x 4 threads, 192 CPUs, completed 2026-03-23 17:41 EDT

\end{document}
